---
title: "Gene Annotations for Nematostella vectensis"
output: github_document
---

```{r echo=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=TRUE, warning=FALSE, message=FALSE)
```

```{r, echo=FALSE}
library(tidyverse)
```


This document combines results from a wide variety of analysis programs to produce a single table with one row per protein.  It assumes that outputs from annotation programs are present in `latest/annotation/`

Raw data includes;

*Protein lengths*
```{r}
# Lengths
#
protein_lengths <- read_tsv("latest/annotation/GCF_000209225.1_ASM20922v1_protein.lengths.txt",col_types = cols(),col_names = c("nemvec_id","protein_length"))
```

*Outputs from blastp and blastx searches against Swissprot*
```{r}
# BLAST
#
bl6_cols <- c("qaccver","saccver","pident","length","mismatch","gapopen","qstart","qend","sstart","send","evalue","bitscore")
blastp <- read_tsv("latest/annotation/GCF_000209225.1_ASM20922v1_protein.blastp.outfmt6",col_names = bl6_cols,col_types = cols()) %>% 
  select(nemvec_id = qaccver, evalue,uniprot_id=saccver) %>% add_column(method = "blastp")
blastx <- read_tsv("latest/annotation/GCF_000209225.1_ASM20922v1_protein.blastx.outfmt6",col_names = bl6_cols,col_types = cols())  %>% 
  select(nemvec_id = qaccver, evalue,uniprot_id=saccver) %>% add_column(method = "blastx") %>% 
  mutate(nemvec_id = str_match(nemvec_id, pattern = "cds_([A-Z]+_[0-9]+.[0-9])")[,2])

blast <- rbind(blastp, blastx) %>% 
  group_by(nemvec_id) %>% 
  top_n(1,desc(evalue)) %>% 
  group_by(nemvec_id, evalue) %>% 
  filter(row_number()==1)

blast_filtered <- blast %>% filter(evalue < 1e-5)

#Uniprot information is a bit special because it is used to amend the blast result with more info. The uniprot table is #therefore joined with the blast results (only) via their uniprot ID's.

uniprot <- readxl::read_excel("latest/annotation/uniprot.xlsx", guess_max = 10000) %>% 
  select(-starts_with("yourlist"))

blast_uniprot <- blast_filtered %>% left_join(uniprot, by=c("uniprot_id"='Entry name'))
```


*SignalP prediction*
```{r}
# SignalP
#
signalp <- read_tsv("latest/annotation/GCF_000209225.1_ASM20922v1_protein.signalp.out",comment = "#", col_names = c("nemvec_id","source","feature","start","end","score","N","A","has_signal")) %>% 
  filter(has_signal=="YES") %>% select(nemvec_id,has_signal)
```

*Pfam domain assignments*
```{r}
# Pfam
#
pfam <- read_tsv("latest/annotation/pfam_tabular.txt",comment = "#",col_names = c("nemvec_id","pfam_id","evalue","description")) %>%
  group_by(nemvec_id) %>% 
  summarise(pfam = paste(unique(pfam_id),collapse=";"))
```

*Interproscan results*
These provide huge amounts of information but we only use the GO term assignments. We therefore transform interpro results to a simple table mapping adi ids to lists of GO terms
```{r}
# InterproScan
# Warns about rows with too few columns.  These can be ignored 
#
interpro <- read_tsv("latest/annotation/nemvec.tsv", guess_max = 10000, 
                     col_names = c("nemvec_id","md5","seqlen","analysis","signature_acc","signature_desc","start","stop","score","status","date","ipr_acc","ipr_ann","ipr_go", "pathways"))

collapse_go <- function(goterms){
  tmp <- goterms %>% str_split("\\|") %>% unlist() 
  tmp <- tmp[!is.na(tmp)]
  paste(tmp,collapse = ";")
}

interpro_go <- interpro %>% group_by(nemvec_id) %>% 
  summarise(ipr_go = collapse_go(ipr_go))
```

Finally all the data is merged into a single table and written to `annotation_table.tsv`
```{r}
  merged <- protein_lengths %>% 
  left_join(blast_uniprot) %>% 
  left_join(signalp,by="nemvec_id") %>% 
  left_join(pfam,by="nemvec_id") %>% 
  left_join(interpro_go)

write_tsv(merged,"latest/annotation_table.tsv")
```

## Summary

As a crude summary we simply look at how many non-NA values exist against key annotation criteria

```{r, echo=FALSE}
merged %>% 
  select(blast=uniprot_id,
         pfam = pfam,
         interpro_go = ipr_go, uniprot_go = `Gene ontology IDs`) %>% 
  gather(key = "method", value = "annotation") %>% 
  mutate(annotation = ifelse(annotation=="",NA,annotation)) %>% 
  na.omit() %>% 
  group_by(method) %>% 
  summarise(num_annotated = n()/33878) %>% 
  ggplot(aes(x=method,y=num_annotated)) + geom_col() + xlab("Annotation") + ylab("Proportion")
```



